---
title: "HW week 11"
author: "Da Qi Ren"
subtitle: 'w203: Statistics for Data Science'

output:
  pdf_document: default
  html_document: default
---
\newcommand{\E}{\text{E}}
\newcommand{\V}{\text{V}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Cov}{\mathbb{\text{Cov}}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r load packages and set options, include=FALSE}
library(tidyverse) 
library(magrittr)
library(knitr)
library(patchwork)
library(ggplot2)
#library(ggpubr)
#library(rstatix)
library(dplyr)
library(nortest)
library(splitstackshape)
library(haven)
library(nhstplot)
library(gplots)
library(kableExtra)
library(readxl)
library(lmtest)


theme_set(theme_bw())
options(tinytex.verbose = TRUE)
```





## Regression analysis of YouTube dataset

You want to explain how much the quality of a video affects the number of views it receives on social media.  **This is a causal question.**

You will use a dataset created by Cheng, Dale and Liu at Simon Fraser University.  It includes observations about 9618 videos shared on YouTube.  Please see [this link](http://netsg.cs.sfu.ca/youtubedata/) for details about how the data was collected.

You will use the following variables:

- views: the number of views by YouTube users.

- rate: the average rating given by users.

- length: the duration of the video in seconds.

You want to use the `rate` variable as a proxy for video quality.  You also include `length` as a control variable.  You estimate the following ols regression:

$$\text{views} =   789 +  2103    \text{ rate} +      3.00 \text{ length} $$


a. Name an omitted variable that you think could induce significant omitted variable bias.  Argue whether the direction of bias is towards zero or away from zero.


```{r}
dat<-read.csv('k.csv', sep='', na.strings=c(""))

dat<-dat %>% 
  select('rate', 'views', 'length', 'ratings') 

nrow(dat)

dat <- dat[!is.na(as.numeric(as.character(dat$views))),]
dat <- dat[!is.na(as.numeric(as.character(dat$length))),]
dat <- dat[!is.na(as.numeric(as.character(dat$ratings))),]
dat <- dat[!is.na(as.numeric(as.character(dat$rate))),]
dat <- na.omit(dat)
dat <- as.numeric(unlist(dat))

summary(dat)
write.csv(dat, "c.csv")


```


b. Provide a story for why there might be a reverse causal pathway (from the number of views to the average rating).  Argue whether the direction of bias is towards zero or away from zero.

```{r}

c <-lm( dat$views ~ dat$rate + dat$length )

summary(c)

```




c. You are considering adding a new variable, `rating`, which represents the total number of ratings.  Explain how this would affect your measurement goal.
```{r}

```

