---
title: "HW week 11"
author: "Da Qi Ren"
subtitle: 'w203: Statistics for Data Science'

output:
  pdf_document: default
  html_document: default
---
\newcommand{\E}{\text{E}}
\newcommand{\V}{\text{V}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Cov}{\mathbb{\text{Cov}}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r load packages and set options, include=FALSE}
library(tidyverse) 
library(magrittr)
library(knitr)
library(patchwork)
library(ggplot2)
library(ggpubr)
library(rstatix)
library(dplyr)
library(nortest)
library(splitstackshape)
library(haven)
library(nhstplot)
library(gplots)
library(kableExtra)
library(readxl)
library(lmtest)
library(stargazer)

theme_set(theme_bw())
options(tinytex.verbose = TRUE)
```





## Regression analysis of YouTube dataset

You want to explain how much the quality of a video affects the number of views it receives on social media.  **This is a causal question.**

You will use a dataset created by Cheng, Dale and Liu at Simon Fraser University.  It includes observations about 9618 videos shared on YouTube.  Please see [this link](http://netsg.cs.sfu.ca/youtubedata/) for details about how the data was collected.

You will use the following variables:

- views: the number of views by YouTube users.

- rate: the average rating given by users.

- length: the duration of the video in seconds.

You want to use the `rate` variable as a proxy for video quality.  You also include `length` as a control variable.  You estimate the following ols regression:

$$\text{views} =   789 +  2103    \text{ rate} +      3.00 \text{ length} $$
a. Name an omitted variable that you think could induce significant omitted variable bias.  Argue whether the direction of bias is towards zero or away from zero.


```{r}
df<-read.csv('videos001.csv')

#df<-df %>% 
#  select('rate', 'views', 'length', 'ratings') 

nrow(df)
df <- na.omit(df)
nrow(df)
summary(df)
 
write.csv(df, "c.csv")
cor(df$rate, df$ratings)
cor(df$ratings, df$length)

#cor(dat$views ~ dat$rate + dat$length + dat$ratings  )
#cor(dat$views ~ dat$rate + dat$length + dat$ratings  )

```

```{r summary-table, eval=FALSE, include=FALSE}

kable(
  df,
  digits = 2,
  caption = 'Cross Tab of Covid-19 infection and the approval status:  rank of -2, -1, 1, 2  represent strong disapproval, disapproval, approval and strong approval, respectively. Note that, we use 0 represent nutral or no responds that are not included in the test. ', 
  booktabs = TRUE, 
)

```




```{r}
k1 <- lm(df$views ~ df$ratings)
k2 <- lm(df$views ~ df$rate + df$length)
k3 <- lm(df$views ~ df$rate + df$length + df$ratings)

stargazer(k1, k2, k3, type = 'text', omit.stat = 'all')

```






```{r}





summary(c2)

cor(df$views, df$ratings)
cor(df$views, df$rate)
cor(df$views, df$length)
cor(df$ratings, df$rate)
cor(df$ratings, df$length)
cor(df$rate, df$length)

```

Discuss of Omitted Variables 






 


b. Provide a story for why there might be a reverse causal pathway (from the number of views to the average rating).  Argue whether the direction of bias is towards zero or away from zero.

```{r}

c3 <- lm(df$rate ~ df$views)
summary(c3)
stargazer(c3, type = 'text' )

 



```




c. You are considering adding a new variable, `rating`, which represents the total number of ratings.  Explain how this would affect your measurement goal.


```{r}
c4 <-lm(df$views ~ df$rate + df$length + df$ratings)
summary(c4)
```




```{r}

# df$logviews = log(df$views) 
# df$lograte =  log(df$rate) 
# df$loglength = log(df$length) 
# df$ratings = log(df$ratings)
# df <- df[is.finite(rowSums(df)),]
# df <- na.omit(df)
# 
# 
# c5 <-lm(df$logviews ~ df$lograte + df$loglength + df$ratings)
# summary(c5)

```

